|

### 2020-02-29 analysis/paper/manuscript.Rmd
Augment introduction following latest review

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. For well-monitored stocks, the challenge has shifted from having sufficient data and information for providing sound advice on stock status, to presenting large quantities of data and output from increasingly complex statistical models in a meaningful way. Traditional formats (e.g. slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018]. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018]. This trend includes fisheries research and many of these tools, and the code used to produce them, are available online (see Appendix A for examples).

Here we aim to demonstrate how interactive visualization tools provide an efficient and effective means of exploring and communicating the ever-expanding array of data inputs and model outputs. We especially highlight the use of interactive visuals inside dynamic documents called "dashboards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. We first demonstrate the application of these tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive dashboards can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In each case, the dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).

@@

We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful. We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar and four anonymous reviewers.

@@

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/) from Cefas
- Benthic Non Native Species Tool (https://openscience.cefas.co.uk/invasive_species/) from Cefas
- Species Dashboard (https://shiny.marine.ie/speciesdash/) from The Marine Institute
- The Stock Book (https://shiny.marine.ie/stockbook/) from The Marine Institute
- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo) from the Fisheries Center, University of British Columbia
- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/) from Fisheries and Oceans Canada, Maritimes Region
- Catchy Data (https://eu.oceana.org/en/catchy-data) from Oceana
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018) from ICES, which was developed to meet an "EU request on dissemination of ICES advice beyond pdf files"

>>>>

Quantitative stock assessment plays a central role in modern fisheries management [@hilborn1992]. Over time, and as new methods are developed, there is an increasing amount of data available to inform stock assessments. These data might be richer information on stock structure using an increasing array of markers and biomarkers, and/or the continued lengthening of population status and catch time series. Concurrently, analytical methods are now able to integrate many data sources into one stock assessment model [@maunder2013]. While these advances have surely improved the advice provided at stock assessments, it has become more difficult to communicate the basis of this advice in a meaningful way. Traditional formats (e.g. slides) for presenting information at stakeholder meetings are often insufficient to convey the richness of information available and the static and sequential nature of these formats can stifle meaningful discussions. A solution to this problem is the effective use of interactive visualization tools [@keena2016]. These are common tools we use every day on a range of web sites, but their use is no longer restricted to web site developers as these tools are being integrated into software commonly used by the research community [@perkel2018]. As a result, an increasing number of scientists across a wide range of disciplines are starting to apply interactive visualization tools to explore and communicate their results [e.g. @jones2016; @yeatman2018; @letcher2018].

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and test custom parameterizations of a Von Bertalanffy Growth model [@catalan2018; see Appendix A for more examples]. While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts. These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.

We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In both cases interactive visuals are nested inside dynamic documents called "dashboards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).

@@

We thank the numerous colleagues and participants of various stakeholder and stock assessment meetings who encouraged us to further develop these interactive visualization tools, and especially those who took the time to make suggestions on how to make them more accessible and useful. We are also grateful for the constructive feedback from Gary Carvalho, Kelli Johnson, Colin Millar, Daniel Goethel and seven anonymous reviewers.

@@

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/); a benthic-focused tool to assist with marine license application and compliance monitoring.
- Species Dashboard (https://shiny.marine.ie/speciesdash/); a tool for exploring biological fisheries data collected in Irish waters
- The Stock Book (https://shiny.marine.ie/stockbook/); a tool for accessing up-to-date information on the status of fish stocks in Irish waters.
- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo); an application for demonstrating the key features of the management strategy evaluation toolkit.
- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/); a tool accessing results from an analysis exploring links between stock indices and ecosystem trends.
- EcoCast Explorer (https://coastwatch.pfeg.noaa.gov/ecocast/explorer.html); an application for displaying results from a dynamic ocean management tool focused on bycatch avoidance [see @hazen2018 for more detail].
- B-VBGM! Bayesian Von Bertalanffy Growth Model tool (https://fishecology.shinyapps.io/B-VBGM/); a supplement Catalán et al. [-@catalan2018] that facilitates the exploration of the model described in the paper.
- Beyond Temperature tool (https://heatherwelch.shinyapps.io/beyond_temperature/); a supplement to McHenry et al. [-@mchenry2019] that allows the exploration of habitat suitability hindcasts and projections by species and model type.
- BRSApp (https://catalinagomez.shinyapps.io/BRSApp/); a supplement to Gomez et al. [-@gomez2016] that allows the exploration of behavioural response severity scores of marine mammals to noise.
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice (developed to meet an "EU request on dissemination of ICES advice beyond pdf files").
